\documentclass[../main-sheet.tex]{subfiles}
\graphicspath{ {../img/} }
\begin{document}
\chapter*{Questions from Previous Years}
\section*{2014-2015 (2017)}
\begin{enumerate}
    \item Marks : $ 4+4+6=14 $
          \begin{enumerate}
              \item Define metric space with example. Let $ d$ be a metric on a set $ M $ and let $ \rho $ be defined by 
              \[
                  \rho(x,y)=\frac{d(x,y)}{1+d(x,y)}
              \]
              show that $ \rho  $ is bounded metric on $ M $.
              \item Define a normed liner space with example. Prove that an Euclidean space is a normed linear space.
              \item What is meant by an inner product space? State and prove Cauchy-Schwarz inequality.
        \end{enumerate}
    \item Marks : $ 4+3+7=14 $
        \begin{enumerate}
            \item For a metric space $ X $, define 
            \begin{enumerate}
                \item neighborhood of a point of $ X $,
                \item limit point of a subset of $ X $,
                \item closed subset of $ X $,
                \item interior point of a subset $ X $.
            \end{enumerate}
            \item Prove that every neighborhood is an open set.
            \item Let $ X $ be a metric space. Prove that the intersection of finite collection of open sets is open. Also give an example to show that intersection of any collection of open set may not be open.
          \end{enumerate}
    \item Marks : $ 5+4+5=14 $
          \begin{enumerate}
              \item Define compact set with example. Show that the set of real numbers is not a compact set.
              \item Explain what is meant by connected set. Is the set of integers a connected set? Support your answer.
              \item If $ \set{K_\alpha} $ is a collection of compact subsets of a metric space $ X $ such that the intersection of every finite sub-collection of $ \set{K_\alpha} $ is non-empty, then prove that $ \cap K_\alpha $ is non-empty.
            \end{enumerate}
    \item Marks : $ 5+4+5=14 $
        \begin{enumerate}
            \item Define continuous function in a metric space. Prove that a function $ f $ of a metric space $ X $ into a metric space $ Y $ is continuous if and only if  $ f^{-1}(V) $ is open $ X $ for every open set $ V $ in $ Y $.
            \item Prove that every continuous image of a connected subset of a metric space is connected.
            \item Show that the set $ \set{x \in \R^n : \norm{x} = 1} $ is compact and connected.
        \end{enumerate}\newpage
    \item Marks : $ 8+6=14 $
        \begin{enumerate}
            \item State and prove the Maximum and Minimum theorem. Hence, verify it for
            \[
                f(x)=\frac{x}{1+x^2}
            \]
            on $ [0, 1] $.
            \item State Heine-Borel Theorem. Prove that every compact subset of a metric space is closed.
        \end{enumerate}
    \item Marks : $ 5+5+4=14 $
        \begin{enumerate}
            \item What is the Intermediate Value Theorem? Can you prove it? How?
            \item Show that there exists a real root of a cube polynomial.
            \item Let $ f: (a,b) \to \R $ be differentiable, and suppose that there is a constant $ M > 0 $ such that $ \abs{f'(x)} \leq M $ for all $ x \in (a,b) $. Show that $ f $ is uniformly continuous on $ (a, b) $.
        \end{enumerate}
    \item Marks : $ 5+4+5=14 $
        \begin{enumerate}
            \item Define pointwise convergent and uniform convergence of a sequence of functions in metric spaces. Let $ f_n:\R\to\R  $ be defined by 
            \[
                f_n(x)=\frac{\sin x}{n}
            \]
            Show that $ f_n\to f=0 $ uniformly as $ n\to \infty $.
            \item Let $ f_n(x)=x^n,\;0\leq x\leq 1 $. Does $ f_n $ converge uniformly? What is the situation for $ 0\leq x<1 $?
            \item Suppose $ f_n(x)=\frac{x^n}{1+x^n} $ for $ x\in[0,2] $. Show that $ \seq{f_n} $ converges pointwise on $ [0,2] $ but that the convergence is not uniformly.
        \end{enumerate}
    \item Marks : $ 7+7=14 $
        \begin{enumerate}
            \item State the Contraction Mapping Principle. Consider the integral equation\\ $ f(x)=a+\int_0^x f(y)xe^{-xy}\dy $. Check directly on which intervals $ [0,r] $ we get a contraction.
            \item Let $ \Omega $ be the set of all invertible linear operator on $ \R^n $. Then prove that $ \Omega $ is an open subset of $ L(R^n) $, the set of all linear transformation on $ \R^n $ and the mapping $ A\to A^{-1} $ is continuous on $ \Omega $.
        \end{enumerate}
\end{enumerate}
\section*{2015-2016 (2018)}
\begin{enumerate}
    \item Marks : $ 5+5+4=14 $
          \begin{enumerate}
              \item Define Euclidean space. Prove that Euclidean space is a metric space.
              \item What do you mean by supremum norm? If $ (V,\seq{\cdot,\cdot}) $ is an inner product space and $ \norm{\cdot} $ is defined for $ v \in V $ by $ \norm{v}=\sqrt{\seq{v,v}} $, then prove that $ \norm{\cdot} $ is a norm on $ V $.
              \item Let $ C([0, 1]) $ be the vector space of all continuous functions defined on $ [0,1] $. For $ f,g \in C([0,1]) $ define $ \seq{f,g}= \int_0^1 f(x) g(x)\dx $. Show that $ \seq{f, g } $ is an inner product and hence $ C([0, 1]) $ is an inner product space.
        \end{enumerate}\newpage
        \item Marks : $ 3+4+7=14 $
        \begin{enumerate}
            \item Define interior, boundary, and closure of a set with examples.
            \item Let $ A \subset \R^n  $ be open and $ B\subset  \R^n $, define $ A+B =\set{x+y\in\R^n\mid x \in A \text{ and } y \in B} $, prove that $ A + B $ is open.
            \item In a metric space, prove that the union of finite number of closed subsets is closed, and the intersection of arbitrary collection of closed subsets is closed. What do you say about the union of an infinite collection of closed sets?
        \end{enumerate}
        \item Marks : $ 6+8=14 $
        \begin{enumerate}
              \item If $ \set{K_\alpha} $ is a collection of compact subsets of a metric space $ X $ such that the intersection of every finite sub-collection of $ \set{K_\alpha} $ is non-empty, then prove that $ \bigcap_\alpha K_\alpha $ is non-empty.
              \item State and prove the Heine-Borel theorem.
        \end{enumerate}
        \item Marks : $ 4+4+6=14 $
          \begin{enumerate}
            \item Define Cauchy sequence in a metric space. Prove that every convergent sequence in a metric space is a Cauchy sequence.
              \item Define diameter of a subset of a metric space. Prove that diam$ \bar{E}= $diam$ E $ where $ \bar{E} $ is the closure of a subset $ E $ of a metric space.
              \item Define complete metric space. Prove that $ \R^k $ is complete.
            \end{enumerate}
    \item Marks : $ 7+7=14 $
    \begin{enumerate}
              \item Define pointwise convergence and uniform convergence of sequence of functions in a metric space. Distinguish between these two sorts of convergences. Show that the sequence of functions
              \[
                  f_k(x)=\begin{cases}
                  0 & x\geq \sfrac{1}{k}\\
                  -kx+1 & 0\leq x < \sfrac{1}{k}
                \end{cases}
              \] 
              converges pointwise on $ [0, 1] $.
              \item Can we differentiate and integrate an infinite series of function? When and how? Examine the uniform convergence of the series
              \[
                  \sum_{n=0}^\infty \frac{(\sin nx)^2}{\pi^2}
                  \]
                  on $ \R $.
        \end{enumerate}
        \item Marks : $ 7+7=14 $
        \begin{enumerate}
              \item Prove that a mapping $ f: X \to Y $ of metric spaces is continuous if and only if $ f^{-1}(V) $ is open in $ X $ for every open set $ V $ in $ Y $.
              \item Define uniformly continuous mapping on a metric space. Prove that every continuous mapping on compact metric spaces is uniformly continuous.
            \end{enumerate}
    \item Marks : $ 7+4+3=14 $
    \begin{enumerate}
        \item State and prove the Maximum-Minimum Theorem. Hence, justify it for $ f(x) = x^3 - x $ on $ [-1, 1] $.
              \item Let $ f: [1,2] \to [0, 3] $ be continuous such that $ f(1) = 0 $ and $ f(2) = 3 $, show that $ f $ has a fixed point.
              \item Let $ f:(0,1]\to\R  $ be defined by $ f(x)=\frac{1}{x} $. Show that $ f $ is uniformly continuous on $ [a,1] $ for $ a>0 $.
        \end{enumerate}
        \item Marks : $ 7+7=14 $
          \begin{enumerate}
            \item State and prove the Contraction Mapping Principle.
            \item Find a function $ f $ such that $ f'(x) = xf (x) $ for $ x $ near $ 0 $ and $ f(0) = 3 $.
        \end{enumerate}
    \end{enumerate}
\section*{2016-2017 (2019)}
    \begin{enumerate}
        \item Marks : $ 4+4+6=14 $
              \begin{enumerate}
                  \item Define metric space with an example. Let $ (M,d) $ be a metric space. Construct a function $ \rho(x,y)=\frac{d(x,y)}{1+d(x,y)}  $. Then show that $ (M, \rho) $ is a metric space and $ \rho(x, y) $ is bounded by 1.
                  \item What is meant by a normed space? If $ (V, \norm{\cdot}) $ is a normed space, then prove that it is a metric space defined by the function $ d(x, y) = \norm{x - y} $.
                  \item State and prove Cauchy Schwarz Inequality. Put the inner product $ \seq{f,g }= \int_0^1 f (x) g(x)\dx  $ on the space of all continuous functions $ C([0,1]) $. Then verify the stated inequality with $ f(x) = x $ and $ g(x) = x^2 $ on $ [0,1] $.
            \end{enumerate}
        \item Marks : $ 5+6+3=14 $
              \begin{enumerate}
                  \item Define neighbourhood of a point in a metric space with example. Prove that neighbourhood of a point in a metric space is an open set.
                  \item Prove that intersection of a finite number of open subsets of a metric space is open. How is about the intersection of an arbitrary family of open sets? Justify your answer.
                  \item Find the interior, exterior, and boundary of $ S = {(x, y): x^2 + y^2 \leq 1} $.
            \end{enumerate}
        \item Marks : $ 5+4+5=14 $
              \begin{enumerate}
                  \item Let $ K \subset Y \subset X $. Then prove that $ K $ is compact relative to X if and only if $ K $ is compact relative to $ Y $.
                  \item Prove that every infinite subset $ E $ of a compact space $ K $ has a limit point in $ K $.
                  \item Prove that every $ k- $cell is compact.
            \end{enumerate}
        \item Marks : $ 7+7=14 $
              \begin{enumerate}
                  \item Define compact set, connected set, and path-connected set with example. Show that $ A = {x \in \R^n : \norm{x} \leq 2} $ is compact and connected.
                  \item State and prove the Heine-Borel Theorem.
            \end{enumerate}
        \item Marks : $ 7+7=14 $
              \begin{enumerate}
                  \item Define continuous function on a metric space. Prove that a mapping $ f:X \to  Y $, where $ X $ and $ Y $ are metric space, is continuous if and only if $ f^{-1}(V) $ is open in $ X $ for every open set $ V $ in $ Y $.
                  \item Define uniformly continuous function on a metric space. Prove that every continuous mapping from a compact metric space into a metric space is uniformly continuous.
            \end{enumerate}
        \item Marks : $ 5+4+5=14 $
              \begin{enumerate}
                  \item State and prove the Intermediate-Value Theorem.
                  Show that f is uniformly continuous on (a, b).
                  \item Show that the equation $ ax^3 + bx^2 + cx + d = 0 $, $ (a \neq 0) $ has at least a real solution.
                  \item Let $ f: (a,b) \to \R $ be differentiable, and suppose that there is a constant $ M > 0 $ such that $ \norm{f'(x)} \leq M $ for all $ x \in (a,b) $. Here, $ a $ or $ b $ may be $ \pm \infty $, and $ f' $ stands for the derivative of $ f $.
            \end{enumerate}
        \item Marks : $ 8+6=14 $
              \begin{enumerate}
                  \item Distinguish between pointwise convergence and uniform convergence for sequence of functions. Show that the sequence $ f_n (x) = \frac{x^n}{(1+x^n)},\,x \in [0,2] $ converges pointwise on $ [0,2] $ but that the convergence is not uniform.
                  \item Prove that every Cauchy sequence in a compact metric space is convergent.
            \end{enumerate}
        \item Marks : $ 7+7=14 $
              \begin{enumerate}
                  \item State and prove the Inverse Function Theorem.
                  \item Define contraction of a metric space. If $ X $ is a complete metric space, and $ \varphi $ is a contraction on $ X $, then prove that there is a unique $ x \in X $ such that $ \varphi(x) = x $. 
            \end{enumerate}
\end{enumerate}
\end{document}